Querying Integrated Ecological Data

### 2010-04-28 docs/algdraft/DataIntegration/DataIntegration.tex
update the data integration draft with more clarified query definition and example.

\documentclass[10pt]{article}
%\title{}
%\author{}
%\date{}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{a4}
\usepackage{color}
\renewcommand\floatpagefraction{0.99}
\renewcommand\topfraction{0.99}
\renewcommand\bottomfraction{0.99}
\renewcommand\textfraction{.05}
\setcounter{totalnumber}{5}

%\setlength{\parindent}{0pt}
\setlength\topmargin{0in}
\setlength\headheight{0in}
\setlength\headsep{0in}
\setlength\textheight{8.5in}
\setlength\textwidth{6.5in}
\setlength\oddsidemargin{0in}
\setlength\evensidemargin{0in}

@@

\title{Data integration}
\author{}
\begin{document}
\maketitle

\section{Materialize Database}

@@

\section{Query rewriting}

What kinds of queries?

@@

IR-style query: 
\begin{itemize}
\item KQ1: Give me the datasets that contain species ``Picea rubens'' observations.
\item KQ2: give me the datasets that have measurements on ``area''
  characteristics. 
\item KQ3: Give me the datasets that contain species ``Picea rubens''
  observations in ``California''. 
\end{itemize}

IR-style summarization query: 
\begin{itemize}
\item SQ1: Give me the dataset that contains at least five distinct ``Picea rubens'' observations.
\item SQ2: Give me the datasets that have measurements with average ``area''
bigger than 5.0 square feet. 
\end{itemize}

@@

How to perform query rewriting? 
m1:

data1 --> table 1
data2 --> table 2 
When we put the data as data tables, we add a unique record id for
each row in the data. This is used to get the unique record id.

Metadata
Table definition

Table column definition

Map annotation information

\begin{example}
KQ1: <concept>: <cond>
Concept can be observation, measurement, characteristic, standard,
unit, etc.

Use query rewriting method: 
For KQ1, give me the datasets that contain species ``Picea rubens''
observations.
concept = observation, cond is species = ``Picea rubens'';

@@

\begin{example}
SQ1: Give me the dataset that contains at least five distinct ``Picea
rubens'' observations.

@@

@@

>>>>

\documentclass[conference]{IEEEtran}

\usepackage{algorithm}
\usepackage{algorithmic}
\renewcommand\floatpagefraction{0.99}
\renewcommand\topfraction{0.99}
\renewcommand\bottomfraction{0.99}
\renewcommand\textfraction{.05}
\setcounter{totalnumber}{5}

@@

\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

\title{Querying Integrated Ecological Data}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Michael Shell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Homer Simpson}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678-2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

\maketitle

\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

% no keywords

\section{Introduction}
In this work, we study the query processing over integrated ecological
data through OBOE model\cite{DBLP:conf/er/BowersMS0}, a conceptual model used to interpret
observation and measurement data.

Contributions of this work:
\begin{itemize}
\item Formalize the queries over observation and measurement query.
\item Propose three methods to evaluate queries. 
\end{itemize}

\section{Related work}

This is related

\section{Data model and query}

In the ecological domain, observation and measurement data are the
core. As a consequence, people are more interested in finding datasets
related to specific observation and measurements.

\subsection{Data model}

We a scientist contributes dataset into the data repository, generally
each dataset is converted to a data table. 
So, the database contains metadata about each dataset, the data table
definition (e.g., attribute/column name, attribute type, etc).

To formalize the senario, we use $D$ to denote the set of data tables. Let $d$ be one data table, 
$d[attr]$ or $d[idx]$ represents the set of values for the attribute
$attr$ or for the $idx$-th attribut.

Generally, the dataset is connected to the OBOE model using
annotations.

\from{HP}{Put the screendump of annotation.}

Map annotation information 
Through the annotation, 
We use $A$ to denote the annotation of one dataset.

\from{HP}{Another way to organize the data is to store them in one big
data table with three columns data\_file, attribute\_name,
attribute\_value. The problem with this organizationo is that the
attribute\_value are different types. }

%Implementation consideration: 
%When we put the data to data tables, we add a unique record id for
%each row in the data. This is used to get the unique record id.

\subsection{Observation and measurement query}

Given dataset in Figure \ref{}.

\begin{itemize}
\item $Q_1$: Give me the datasets that contain species ``Picea rubens'' observations.
%$Q_2$: Give me the datasets that have measurements on ``area''  characteristics. 
\item $Q_2$: Give me the datasets that contain species ``Picea rubens''
  observations in ``California''. 
\item $Q_3$: Give me the dataset that contains at least five distinct ``Picea rubens'' observations.
\item $Q_4$: Give me the datasets that have measurements with average ``area''
bigger than 5.0 square feet. 
\end{itemize}

We formalize the queries as follows.

\begin{definition}
KQ1: $<concept>:<cond>$
Here, <concept> can be observation, measurement, characteristic, standard,
unit, etc.. 
<cond> is in the form of $func([distinct]~attr)~op~value$, here $op$ can be some
algorithmmic operators $=$, $\neq$, $>$, or $<$. 
$func$ are aggregation functions including sum, average, min and max.

The result is a set of $d$, $\{d|d\in D \wedge (value~op~dv = true |dv\in d[attr])\}$.
\end{definition}

For each of such result, we denote that $d~s.t.~KQ$ meaning that
dataset $d$ satisfies the query.

This is the baseline keyword style query. 
The following two queries are the basic queries.

\begin{definition}
KQ2: $Q_1~context(Q_1.concept,Q_2)~Q_2 $, here $Q_i$ is a basic query
in the form of $<concept>:<cond>$. 
\end{definition}

\begin{example}
The given example queries can be represented in the formal query
format. 
\begin{itemize}
\item $Q_1 = Observation: Species='Picea rubens'$
\item $Q_2 = <Observation_1: Species='Picea rubens'>
  IN(Observation_1,Observation_2) <Observation_2:
  measurement_{unknown}=`California'>$
\item $Q_3 = Observation: Species='Picea rubens' \wedge
  count(distinct~Species)\geq 5$ \from{HP}{TODO}
\item $Q_4$ 
\end{itemize}

\end{example}

\section{Querying annotated and integrated ecological data}

All the datasets can be imported to a RDBMS for

To answer such queries, a naive method is to

\subsection{Materialized database}

@@

\subsection{Query rewriting}

@@

@@

\begin{example}
Take KQ1 as example. 
concept = observation, cond is {\em species = ``Picea rubens''};

@@

\begin{example}
We use $SQ_1$ as an example.

@@

Use materialized database: 
Analysis: ``Area'' is the context of some observation.
Area --> mlabel, observation type

From the context observation lable, find the observation labels, \\
select olabel\\
from context\_relationship\\
where context\_olabel = 'olabel';

Let olabel\_set  be the set containing all the observation type
labels that use ``Area'' as context.

SELECT oi.record\_id \\
FROM observation\_instance oi, measurement\_instance mi, 
observation\_instance coi, measurement\_instance cmi,context\_instance ci\\
WHERE (mi.oid=ci.oid AND ci.oid = oi.oid) and (cmi.oid=coi.oid AND ci.context\_oid =
coi.oid) AND coi.otype $\in$ olabel\_set 
AND cmi.mlabel=`Area'
GROUP BY cmi.oid
HAVING avg(cmi.mvalue)>5.0in;

@@

\section{Conclusion}

%\bibliographystyle{abbrv}
%\bibliography{model}
%\bibliographystyle{IEEEtran}
%\bibliography{IEEEabrv,DataIntegration}

%\begin{thebibliography}{1}
%\bibliographystyle{abbrv}
%\bibliography{model}
%\end{thebibliography}
